Title here

### 2016-09-19 portfolio/text/ms.Rmd
updated ms

<!-- Fishing is inherently risky and we have long-sought regulations and fishing behaviors that reduce this risk. -->
<!-- The increasing the diversity of species fished has been proposed to reduce risk for fishermen -->

Individuals relying on natural resources as their primary source of income
often experience high levels of income variability
due to the combined effects of environmental, biological, and price variation.
Commercial fishers are a prime example and
recent work has sought to identify regulations and fishing behaviours
that can reduce this variability.
One candidate has been that fishing higher levels of species diversity
decreases income variability.
However, it is unknown if this species-diversity effect
on income variability holds
across multiple approaches to diversifying catch,
at the level of individual permit holders,
and across different fishing strategies.
Here, we simultaneously estimate the effects of three
methods of diversification for Alaskan fishery permit holders
on revenue and revenue variability:
(1) owning multiple permits,
(2) owning permits that inherently fish higher species diversity,
and (3) increasing species diversity within the constraints of a permit set.
We control for changes in effort and other
latent effects through time.
We find that
fishing an additional permit is associated with higher revenue
and lower variability (e.g.,
`r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] lower
expected standard deviation
for halibut plus another permit vs. single halibut permits);
fishing permits with higher species diversity reduces variability
(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
increase in mean effective species diversity);
but that increasing species diversity within
the constraints of a permit set has a fishery-dependent effect on revenue and
is usually (`r g2_dens`% probability)
associated with increased uncertainty in revenue.
At the same time, Alaskan commercial fisheries
have been increasingly focussed on
single-permit (often salmon) fishing strategies
with single-permit fishers earning
more than multiple-permit fishers
for the first time in 2013.
Our results suggest that the benefit of species diversity
for revenue variability
is mostly an across-permit effect, not
a benefit that can be easily harnessed by an individual permit holder.
Effective methods of diversification ---
purchasing additional permits or switching permits ---
are costly and often limited by regulations such as limited entry
programs and individual transferable quotas (ITQs).
Take home ...

@@

- there is often limited room for intentional diversification within a permit
- in fact, specializing can be profitable within some fisheries

@@

Natural resources such as forests, land animals, and fishes
provide a primary source of income for millions of people worldwide [REFS].
However, environmental,
biological, and price variation
combined with regulatory changes
and human-caused environmental disasters
generate high levels of income variability and risk
for these individuals.
For example, ...
And EVOS example

@@

@@

@@

Despite this existing work,
key questions remain on the relationship between species diversity and
fishing revenue variability.
First, the effect of changing species diversity
for individual permit holders on a year-by-year basis is unknown ---
existing approaches have worked at the vessel or community level
and condensed time series into single metrics of variability or risk.
Second, the effect of changing species diversity
on revenue itself is unknown ---
existing approaches have focused solely on variability or risk.
Third, it is unknown if the approach to
diversifying species fished affects the relationship between
species diversity and revenue variability ---
a fisheries permit holder can purchase
an additional permit to target other species,
pick a permit that inherently targets many species,
or increase species diversity
within the constraints of his or her existing permits.
Many permit holders may be limited to the last approach
given that the first two may be costly and limited by regulations,
but it is unclear if the benefits and costs are the same.
Here, we address these questions by
simultaneously estimating the effects of
these three approaches to species diversification
for Alaskan fishery permit holders on
annual revenue and revenue variability.
Our model makes use of the entire time series,
controls for effort and latent time effects,
and allows the effect of species diversity
to vary by fishing strategy (sets of similar permits)
in a hierarchical model.

@@

We combined similar permits into permit groupings 
and identified sets of 
permit groupings held by an individual 
in a specific year as 'strategies' (Supporting Information). 
For example, we combined permits that were the same except for vessel size,
and we combined a number of permits that were targeting a single species 
and differed only in region fished.
Therefore, someone fishing herring roe in Kodiak
and someone fishing herring roe in Alaska Peninsula
were both considered to have a 'herring roe' permit group.
For salmon permits, 
we grouped permits using the same gear type across regions
for groups that shared similar species composition. 
After this process of combining permits,
we were left with `r n_grouped_permits`
unique fishing permit groups (Table S\ref{tab:strategies}),
which we aggregated
within each person-year combination 
to form permit strategies.
For example, if someone fished halibut
and sablefish in the same year,
their strategy for that year would be `halibut-sablefish'.
Specific details on our strategy definitions are available in 
the Supporting Information.

Our models focused on species diversity
as a possible predictor of revenue and revenue variability
while controlling for effort.
We calculated effective species diversity for
permit holder $i$ and time (year) $t$ as the inverse of
Simpson's diversity index $\lambda$ [@simpson1949]
weighted by revenue $R$ ('species diversity'):

@@

The basic form of our model was a 
hierarchical linear regression with Gaussian errors
predicting $\log$ revenue, $\log(R_{i,t})$, for
each fisher $i$ and time $t$:

@@

\noindent
where $S_{i,t}$ is the percent change in species diversity,
$\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$,
$D_{i,t}$ is the percent change in days fished,
$\log \left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and
$I_{i,t}$ is an indicator variable that takes a value $0$
if $\mathrm{div}_{i,t} \ge \mathrm{div}_{i,t-1}$
(species diversity increasing; generalizing)
and takes a value $1$ if
$\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$
(species diversity decreasing; specializing).

The term $\log(R_{i,t-1}$) is treated as an offset term:
it is added on the right side of the equation
to essentially model percent change in revenue,
i.e., $\log(R_{i,t}) - \log(R_{i,t-1}) = \log(R_{i,t}/R_{i,t-1})$.
While this approach required filtering the data
to remove person-year combinations without a subsequent year of revenue,
it has a number of advantages:
it eliminates having to deal with the non-stationary
time series and eliminates having to
estimate thousands of people-specific intercepts.

The $\beta$'s represent estimated coefficients, some of which are
allowed to vary.
The intercepts $\beta_{0,j,t}$ are allowed to vary by
strategy-year combinations,
are centered on zero, and are constrained by a normal distribution:

>>>>

Title should emphasize individuals, e.g. "Benefits of diversification and risks of changing strategies for individual fishermen/fishers"
<!-- Fishing is inherently risky and we have long-sought regulations and fishing behaviors that reduce this risk. -->
<!-- The increasing the diversity of species fished has been proposed to reduce risk for fishermen -->

Individuals relying on natural resources as their primary source of income
often experience high levels of income variability
due to the combined effects of environmental, biological, and price variation.
Commercial fishers are a prime example and
recent work has sought to identify regulations and fishing behaviours
that can reduce this variability.
One candidate has been that fishing higher levels of species diversity
decreases income variability.
However, it is unknown if this species-diversity effect
on income variability holds
across multiple approaches to diversifying catch,
at the level of individual permit holders,
and across different fishing strategies.
Here, we simultaneously estimate the effects of three
methods of diversification for individual participants in Alaskan fisheries
on revenue and revenue variability:
(1) diversifying by owning multiple permits,
(2) diversifying by owning permits that inherently increase species diversity,
and (3) increasing species diversity within the constraints of a permit set.
We control for changes in effort and other
latent effects through time.
We find that
fishing an additional permit is associated with higher revenue
and lower variability (e.g.,
`r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] lower
expected standard deviation
for halibut plus another permit vs. single halibut permits);
fishing permits with higher species diversity reduces variability
(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
increase in mean effective species diversity);
but that increasing species diversity within
the constraints of a permit set has a fishery-dependent effect on revenue and
is usually (`r g2_dens`% probability)
associated with increased uncertainty in revenue.
At the same time, Alaskan commercial fisheries
have been increasingly focussed on
single-permit (often salmon) fishing strategies
with single-permit fishers earning
more than multiple-permit fishers
for the first time in 2013.
Our results suggest that the benefit of species diversity
for revenue variability
is mostly an across-permit effect, not
a benefit that can be easily harnessed by an individual permit holder.
Effective methods of diversification ---
purchasing additional permits or switching permits ---
are costly and often limited by regulations such as limited entry
programs and individual transferable quotas (ITQs).
Take home ...

@@

- there is often limited room for intentional diversification within a permit
- in fact, specializing can be profitable within some fisheries
- analog here in farming literature is that diversification for small scale farmers generally good, except when they specialize on high demand/high value crops (tobacco, cocoa). Seems like benefits of diversification are also smaller for large-scale industrial farmers.

@@

Natural resources such as forests, land animals, and fishes
provide a primary source of income for millions of people worldwide [REFS].
However, environmental,
biological, and price variation
combined with regulatory changes
and human-caused environmental disasters
generate high levels of income variability and risk
for these individuals.
For example, drought and flooding are major sources of risk to agricultural food security (http://rstb.royalsocietypublishing.org/content/360/1463/2139.short),

And EVOS example

@@

- crop rotation has been proposed as metric for buffering against short term variation and long term trends (like climate change)

BUT in some situations, external forces (prices, demand) may discourage the adoption of these diverse strategies and help promote specialization, 
http://link.springer.com/article/10.1007/s10584-004-0710-z

@@

Most papers on fisheries have adopted the approach similar to Kasperski et al. and used data reported to the level of vessels or higher. This analysis is unique in that we're able to separate individual people from vessels (enabling us to track people that switch fishery sectors over time, or change strategies and utilize multiple vessels)

@@

Despite this existing work,
key questions remain on the relationship between diversty of fish catches and
fishing revenue variability.
First, the effect of changing species diversity
for individual permit holders on a year-by-year basis is unknown ---
existing approaches have worked at the vessel or community level
and condensed time series into single metrics of variability or risk.
Second, the effect of changing species diversity
on revenue itself is unknown ---
existing approaches have focused solely on variability or risk.
Third, it is unknown if the approach to
diversifying species fished affects the relationship between
species diversity and revenue variability ---
a fisheries permit holder can purchase
an additional permit to target other species,
pick a permit that inherently targets many species,
or increase species diversity
within the constraints of his or her existing permits.
Many permit holders may be limited to the last approach
given that the first two may be costly and limited by regulations,
but it is unclear if the benefits and costs are the same.
Here, we address these questions by
simultaneously estimating the effects of
these three approaches to species diversification
for Alaskan fishery permit holders on
annual revenue and revenue variability.

I think we need a little here on the AK data and why that's both important and generalizable to fisheries around the world. 
Fisheries in Alaska are the highest valued fisheries in N. America
with landings valued at XX $ billion annually. There are a number of fishing
strategies where both specialization and diversification are extremely profitable. There's a number of fishing permits that constrain the potential diversity of species landed (e.g. herring, halibut, sablefish, or crab permits that allow the user to only land that species) versus broader 'finfish' permits that allow users to land a wide variety of groundfish spp. 
- perhaps we also need to have a sentence or two on the implict constraints of geography and gear / vessel size on species diversification

Our model makes use of the entire time series,
controls for effort and latent time effects,
and allows the effect of species diversity
to vary by fishing strategy (sets of similar permits)
in a hierarchical model.

@@

We combined similar permits into permit groupings 
and identified sets of 
permit groupings held by an individual 
in a specific year as 'strategies' (Supporting Information). 
For example, we combined permits that were the same except for vessel size (e.g. a longliner fishing sablefish on a vessel under 60ft was considered to have the same strategy as a longliner fishing sablefish on a vessel over 60ft),
and we combined a number of permits that were targeting a single species 
and differed only in region fished.
Therefore, someone fishing herring roe in Kodiak
and someone fishing herring roe in Alaska Peninsula
were both considered to have a 'herring roe' permit group.
For salmon permits, 
we grouped permits using the same gear type across regions
for groups that shared similar species composition. 
After this process of combining permits,
we were left with `r n_grouped_permits`
unique fishing permit groups (Table S\ref{tab:strategies}),
which we aggregated
within each person-year combination 
to form permit strategies.
For example, if someone fished halibut
and sablefish in the same year,
their strategy for that year would be `halibut-sablefish'.
Specific details on our strategy definitions are available in 
the Supporting Information.

Our models focused on species diversity
as a possible predictor of revenue and revenue variability
while controlling for effort.
We calculated effective species diversity for
permit holder $i$ and time (year) $t$ as the inverse of
Simpson's diversity index (Jost 2006; also referred to in economics as the Herfindahl-Hirschman Index) $\lambda$ [@simpson1949]
weighted by revenue $R$ ('species diversity') (also say we did this for consistency with previous analyses,e.g. Kasperski and Holland):

@@

The general framework of our model is known as a variance function regression model, which separately models change in the mean and variance as a function of covariates (Davidian and Carrol 1987 http://amstat.tandfonline.com/doi/abs/10.1080/01621459.1987.10478543; Hall and Carrol 1989, http://www.jstor.org/stable/2345837; ). Where traditional linear models or hierarchical linear models generally assume constant variance, the variance function regression model allows variance to be heteroscedastic. For application to economic data on individual fishing portfolios, this type of model allows certain strategies and levels of diversification to be more or less variable than others.

We extended the basic variance function regression to a hierarchical 
Bayesian variance function regression model with Gaussian errors
predicting $\log$ revenue, $\log(R_{i,t})$, for
each fisher $i$ and time $t$:

@@

\noindent
where $S_{i,t}$ is the percent change in species diversity,
$\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$,
$D_{i,t}$ is the percent change in days fished,
$\log \left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and
$I_{i,t}$ is an indicator variable that takes a value $0$
if $\mathrm{div}_{i,t} \ge \mathrm{div}_{i,t-1}$
(species diversity increasing; generalizing)
and takes a value $1$ if
$\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$
(species diversity decreasing; specializing). Estimating a separate
coefficient for the effect of increasing diversity and decreasing
diversity allows the costs and benefits of these changes to be asymmetric.

The term $\log(R_{i,t-1}$) is treated as an offset,
to essentially model percent change in revenue,
i.e., $\log(R_{i,t}) - \log(R_{i,t-1}) = \log(R_{i,t}/R_{i,t-1})$.
This formulation is necessary because of the non-stationary trends in 
revenue (resulting from experience or other unmodeled factors). While this approach required filtering the data
to remove person-year combinations without a subsequent year of revenue,
it has a number of advantages:
it eliminates having to deal with the non-stationary
time series and eliminates having to
estimate thousands of people-specific intercepts.

The $\beta$'s represent estimated coefficients, some of which are allowed to vary as random effect terms.
The intercepts $\beta_{0,j,t}$ are allowed to vary by
strategy-year combinations,
are centered on zero, and are constrained by a normal distribution:
